### 2024-08-08 law-firms-names/Paper.Rmd
Various minor improvements in paper text, code,  style, and refs, #5

---
title: "Spatial Analysis of Russian Law Firms Naming"
author: "Pavel Syomin¹"
output:
  word_document:
    reference_doc: "journal-article/reference.docx"
bibliography: "journal-article/refs.bib"
csl: "journal-article/apa.csl"
link-bibliography: False  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE, 
  echo = FALSE, 
  fig.width = 4.96, 
  fig.id = FALSE,
  dpi = 1200)
IS_PAPER <- TRUE # used in script to disable saving of names and samples
```

```{r imports, cache=TRUE, include=FALSE}
library(here)
source(here("journal-article", "Script.R"))
```

<div custom-style="Affiliation">

¹Perm State University, Geographical faculty, Department of social and economic geography, Perm, Russia; e-mail: ntsp\@ya.ru

</div>

<div custom-style="Abstract">

**Abstract:** The present study analyzes law firm naming in Russia, with a particular focus on the geographical aspects of this phenomenon. The primary objective of this research is to ascertain the impact of the spatial relations between the country's regions, specifically those of a neighboring, proximity, and belonging to a parent-level subdivision, on the naming of law firms. To conduct the analysis, data regarding law firms is extracted from the open dumps of the state registry of small and medium-sized businesses. The names of the firms are transformed into numerical vectors, subsequently clustered with the KMeans method, and then manually grouped to identify naming strategies. Numerical representations of law firm names within each region are averaged to get region embeddings, which are then used to compare the names of law firms in different regions. The research findings indicate that regions situated closer to one another or sharing a common border tend to adopt slightly more similar law firm names. Furthermore, regions that are part of the same economic region also display a greater degree of similarity in law firm naming. Three primary strategies of law firm naming are “Form”, “Law“, and “Service”. Volga and North Caucasus economic regions are characterized by the predominance of the “Service” strategy, whereas the Volga-Vyatka economic region exhibits a “Law” strategy, and the remaining territories demonstrate a combination of naming strategies. The results of the study lend support to the widely used classification of economic regions in Russia. The findings also underscore the impact of neighboring, proximity, and interconnectivity on the economic and cultural attributes of the regions. The administrative open data processing techniques, as well as the methods of working with firm names based on neural networks proposed in this paper, may be used by other economic geographers for their own research purposes.

</div>

<div custom-style="Keywords">

**Keywords:** firm names; law firms; text embeddings; economic regions, text similarity.

</div>

# 1. Introduction

The geography of services, also known as the geography of the service sector, is a crucial component of geographical studies in Russia. This field has a history spanning several decades, with origins dating back to the 1960s. The geography of services has two distinct dimensions: economic and social. The former is primarily concerned with the business aspects of the services, while the latter addresses the implications for human well-being arising from the geographical organization of services [@Tkachenko2016]. The existing scientific literature on the geography of services in Russia covers a broad range of topics, including education [@Shulgina2019], car rental [@Zyrianova2018], catering [@Preobrazhenskiy2018], medical care [@Kazantcev2022; @Panin2023; @Nikoloski2024], and internet access [@Nagirnaya2015; @Zemtsov2022]. However, some areas remain underexplored, with legal services being one such example [@Tkachenko2016; @Moiseeva2016].

Despite the lack of interest from geographers, the spatial aspects of the activity of law firms are worth investigating for a number of reasons. Firstly, legal services can be categorized as business or producer services, and this category of services is important for economic development, on both a national and regional scale [@Hansen1993; @Faulconbridge2019]. Secondly, despite their comparatively minor direct economic impact, these services play a significant role as intermediaries and facilitators, contributing to the establishment and maintenance of the rule of law, social order, stability, and predictability of social relations. Thirdly, for the general public, access to legal aid is vital for the protection of human rights and the attainment of social justice. In this regard, the spatial aspects of law services are on par with the spatial issues of medical care [@Blacksell1988; @Blacksell1990].

The study of firm names — which are sometimes also referred to as company names, brand names, or simply “brands” — is a relatively uncommon yet potentially fruitful approach to the geographical analysis of services and related problems. A firm name is designed to distinguish the business from competitors, to highlight the sector or geographical area in which it operates, and to convey information about the services offered by a provider and the associated benefits. A brand is therefore capable of transferring a substantial amount of information in a brief textual format. Such information may prove useful in uncovering the city's business landscape and consumer perceptions of a particular city with its distinctive services [@Raento2001]. It may also help to analyze trends common to a specific sector in a geographic region [@Marjavaara2022]. The incorporation of place names (toponyms) in brands influences the business development and consumer attitudes [@Peluso2012; @Zhang2014; @Erdogan2016; @Noorda2019]. Company names encapsulate a multitude of intertwined properties relevant to the economy and culture of a given location. This synthesis makes them suitable for the identification of regions, especially in the case of vernacular regions [@Reed1976; @Liesch2015; @Vukosav2016; @Holtkamp2018].

It seems reasonable to posit that the names of Russian law firms reflect some economic and, to a lesser extent, cultural characteristics of the places where these firms are located. The substantial number of such firms (tens of thousands according to @Moiseeva2016) appears to be sufficient for a comprehensive and nuanced representation of areas and territories. It can be hypothesized that there is a geographical variation in the naming of law firms across the country, which is contingent upon a number of factors. The geographical proximity and common borders, for example, should lead to more homogeneous naming practices. Similarly, economic links between places and territories are also expected to reduce the diversity of law firm names. In addition, it may be hypothesized that there are several law firm naming strategies, or typical models of brand construction, that potentially vary across the country. The primary aim of this study is to test these hypotheses.

# 2. Materials and methods

## 2.1. Data

The issue of obtaining information about firm names is a challenging one. All Russian organizations are registered in the state registry operated by the Federal Tax Service (FTS of Russia). An extract from this registry about a particular company can be searched and downloaded online by any person free of charge. However, a considerable sum of money must be paid and a written request submitted to the FTS of Russia to purchase a complete copy of the registry. Commercial companies offer data re-cell services, but the prices are similarly high.

To address these challenges, a unified state registry of small and medium-sized businesses (SMB registry) may be used. The registry has been maintained since 2016 in accordance with Article 4.1 of the Federal Law of 24 July 2007 No. 209-FZ “On the development of small and medium-sized businesses in the Russian Federation”. All companies and individual entrepreneurs with a yearly revenue below 2 billion roubles and an average yearly number of employees below 250 are automatically included in the SMB registry, without any action being required on their part. The registry is fully downloadable at the FTS open data portal at no cost. Additionally, the portal provides a monthly record of updates made since the registry's inception. To sum it up, the SMB registry can be regarded as the most comprehensive free data source offering information on Russian business entities, both organizations and individuals.

The downloadable version of the registry includes, in particular, firm names, addresses of registration (up to a settlement), and codes of activity kinds according to the National Classification of Economic Activities (NCEA). It should be noted that the registry does not include large firms, resulting in some data being absent. Nevertheless, the majority of commercial law firms in Russia (in excess of 75%) are classified as small and medium-sized businesses and thus are included in the registry. Consequently, the completeness of information is relatively high, and it can be used to draw reliable conclusions.

## 2.2. Methods

### 2.2.1. Law firms selection criteria

In this study, the term “law firm” is defined as a commercial organization registered in Russia in accordance with the Civil Code of Russia, with a main activity code equal to 69.10 (activity in the area of law) according to NCEA. The aforementioned definition excludes individual entrepreneurs, foreign companies, non-profit organizations, advocacy firms, and notaries. The selection technique based on the main activity code is derived from that proposed by @Moiseeva2016. Individual entrepreneurs are excluded from the analysis because they typically operate under their personal names, and even if they have a brand, it is not listed in the SMB registry. Other types of law firms are not included in the analysis primarily due to data availability issues.

The analysis covers a period of six years, from 2016 to 2021. The initial point in time is defined by the commencement of SMB registry data availability. Although data for 2022 and subsequent years are available, they have not been studied as the ongoing war may have had a significant impact on both the economy and the dissemination of information by government bodies. The purpose of this research is to provide conclusions about the space and names under “normal” conditions, and it would therefore be unwise to consider data collected during the period when the circumstances are away from normal. Organizations with a lifespan of less than three years have been omitted from the research, which are likely to be dummy companies.

### 2.2.2. Firm data acquisition

The processing of SMB registry data is conducted in a series of stages. In the initial phase of the research process, all source archives are downloaded. Each source archive contains a full copy of the registry data for the specified publication date, with a new archive becoming available on a monthly basis. Subsequently, data pertaining to law firms is extracted from the archives and stored in CSV files. Each CSV file contains data from one source archive. Subsequently, all CSV files are consolidated into a single table, the duplicated records are removed, and the remaining records are augmented with start and end temporal markers. Subsequently, the addresses are normalized to resolve the inconsistencies in location names and geocoded to add geographical coordinates. Finally, the resulting table is transformed into a panel view, with each row representing a single law firm at a given year. All these stages are implemented in an open-source Python command-line tool.

### 2.2.3. Firm name embeddings retrieval

The Yandex Foundation Models cloud service is used to transform the firm names into text vectors, or embeddings. The “text-search-query” model is selected because its purpose, as stated in the documentation, is to vectorize short texts. Prior to vectorization, the firm names are preprocessed to remove indications of organizational form such as “limited liability company”. This is done because such indications are common to the majority of names and thus have little value. The results of the queries to the Yandex API are stored separately for reuse, given that API calls are paid and take time to be completed.

### 2.2.4. Region embeddings

The embeddings of firm names are grouped by regions and averaged. In this paper, these average vectors are referred to as region embeddings or region vectors. Russia is comprised of 83 internationally recognized regions (constituent entities of the Russian Federation), hence 83 region vectors are calculated. Each vector represents the combined meaning associated with the names of law firm names within a specific region. The mathematical similarity of region vectors is analogous to the similarity of law firm names in regions.

Regions vectors were introduced in this study to reduce computational complexity. With thousands of law firms, the mathematical operations between individual names would otherwise have been unduly time-consuming. Furthermore, the combined vectors improve the interpretation of the results. A region is a prevalent subdivision used in the studies of Russian economic geography. Therefore, a single sequence of numbers for a single region is more readily understandable than a big table of individual name vectors.

### 2.2.5. Naming strategies identification

Law firm names are clustered using the KMeans technique. A matrix of features for clustering is constructed from the name embeddings. As there are 256 items in each vector, the feature matrix has 256 columns. A target number of clusters is determined with the elbow method and ultimately set to 50, with the understanding that further manual processing of the clusterization results will be necessary. Each name is assigned a cluster label between 1 and 50.

The resulting clusters are subsequently subjected to manual grouping. A random sample of 20 names is selected from each cluster. With 50 clusters and 20 sampled names allocated in each, the total size of the dataset for manual processing is 1, 000. The cluster samples are carefully examined, and a concise description is provided for each, e.g. “Names with the word “консалт/консалтинг” (consult/consulting)”. After this, the clusters are examined once more and assigned to one or more of the four groups: “Form”, “Law”, “Service”, and “Misc”. The mapping of clusters and groups (e.g. cluster 1 — “Law”) is employed to link all the firm names to the appropriate groups. The concept of naming strategies is derived from the works of @Arora2015 and @BassolsiGardella2023. However, the specific groups differ from those used in their respective papers and are drawn from the data-driven analysis.

Each group except for the “Misc” category represents a specific naming strategy. The “Form” strategy encompasses names that are based on a word or words denoting a kind of business entity, including “company”, “firm”, “group”, “partnership”, “center”, “alliance”, and “corporation”. The “Law” strategy emphasizes the firm's affiliation with the legal services market, employing terms such as “legal”, “juridical”, “lex”, “law”, and “lawyer”. Names belonging to the “Service” strategy may indicate a particular service provided by a law firm (“advisor”, “consulting”, “protection”, “support”, “warrant”), describe the sectors of businesses the firm is primarily engaged with (“technology”, “development”, “house management”, “accounting”), or mention services in general. The “Misc” category is a catch-all group for names that do not clearly align with any other category.

### 2.2.6. Analysis of firm names similarity

The similarity of law firm names between regions is calculated as a Euclidean distance between the corresponding region vectors. Henceforth, this will be referred to as a semantic distance between regions. A neighborhood of regions is determined on the basis of the existence of a common border between them, thus allowing for two regions to either be neighboring or not. A geographic distance between regions is calculated as a linear distance between the centroids of the region polygons. To provide a proxy of economic links between regions, the belonging to the same economic region is utilized. According to @Leyzerovich2010, Russia is comprised of 11 economic regions, which themselves consist of a number of regions. The defining characteristics of an economic region are its internal unity, distinction from the other economic regions, and the economic links that exist between constituting regions, cities, and enterprises within them.

To estimate the effect of neighboring, regions are divided into pairs of neighbors and non-neighbors. Semantic distances are then calculated between the regions in each pair within the two groups. A median semantic distance is calculated for each group. It is expected that the median semantic distance between neighboring regions would be less than that between regions that do not share a border.

A similar technique is used to ascertain whether geographical distance impacts semantic distance. Pairwise semantic and geographical distances between all regions are calculated, and the correlation between geographical distance and semantic distance is estimated. It is expected that the coefficient of correlation will be positive.

To analyze the role of economic links, the pairwise semantic distances between regions are split into two groups: distances between regions within an economic region (within-distances) and distances between regions belonging to an economic region and all the other regions (outside-distances). The median of both within-distances and outside-distances is calculated for each economic region. It is expected that semantic distances between regions within an economic region would be smaller than those between regions belonging to different economic regions.

### 2.2.7. Analysis of naming strategies

A dominant naming strategy is ascribed to each region, defined as the naming strategy used by the majority number of law firms within that region. Dominant naming strategies are then mapped and analyzed visually to ascertain whether any geographical patterns emerge. In addition, the boundaries of economic regions are illustrated on the same map, and the dominant naming strategies of regions within each economic region are examined. No a priori hypotheses are formulated regarding the geographical patterns. As for naming strategies within economic regions, it is expected that they will be similar.

# 3. Results and discussion

## 3.1. Law firm names in neighboring regions are more similar

The semantic distance between neighboring regions in comparison to the semantic distance between regions that do not have a common border is illustrated in Figure 1. Although the difference in medians is relatively minor, it is nevertheless present, and the overall distribution of distances between non-neighboring regions is shifted towards the area of higher values. Therefore, neighboring regions have more similar names of law firms than regions that do not border one another.

```{r fig-1-neighbors, fig.cap="**Figure 1.** Semantic distances between neighboring and non-neighboring regions", fig.height=3}
neighbors_plot
```

## 3.2. Greater geographical distance reduces the similarity of law firm names

The relationship between geographical and semantic distances is shown in Figure 2. The correlation coefficient is relatively low but positive. Thus, regions that are situated close to each other tend to have a bit more similar names of law firms even in the absence of a common border between them.

```{r fig-2-distance-plot, fig.cap="**Figure 2.** Relationship between geographical and semantic distances between regions", fig.height=3}
distance_plot
```

## 3.3. Law firm names inside an economic region are more homogeneous

Within-distances and outside-distances by economic regions, as outlined in the methods section, are depicted in Figure 3. In some instances, the semantic distance between regions that are part of the same economic region is slightly less than the semantic distance between these regions and regions of other economic regions. However, this pattern does not apply to at least three economic regions: in Volga, North, and North Caucasus, median within-distance is equal to or greater than median outside-distance. Consequently, the results of this test are mixed: while links between regions constituting an economic region usually slightly increase the similarity of law firm names, there are several exceptions to this rule. A greater within-distance compared to outside-distance may indicate internal heterogeneity within an economic region, thereby suggesting the need for a classification update.

```{r fig-3-economic-regions-distance, fig.cap="**Figure 3.** Semantic distances between regions within and outside economic regions", fig.height=5}
economic_regions_distances
```

## 3.4. Geographical patterns of naming strategies are unclear

The map in Figure 4 depicts a dominant naming strategy in regions of Russia. As “Form” naming is used by a minority of law firms across all regions, it is not shown. European South of Russia and Povolzhye are characterized by a prevalence of the “Service” strategy. In other parts of the country, an irregular mixture of both “Law” and “Service” naming strategies is observed.

```{r fig-4-naming-strategies, fig.cap="**Figure 4.** Map of naming strategies"}
naming_strategies
```

## 3.5. Naming strategies within economic regions are inconsistent

As seen from Figure 4, few economic regions are characterized by a uniform naming strategy. In the Volga-Vyatka economic region, the “Law” strategy is the most prevalent among all the regions, with the exception of the Republic of Mari El. In Volga and North Caucasus economic regions, service-based strategy is the prevailing approach in the majority of regions. The remaining economic regions exhibit a combination of law firms naming strategies.

# 4. Conclusion

The findings of the study are somewhat contradictory. On the one hand, several hypotheses have been proven. These include the following: law firm names in neighboring regions are more similar than law firm names in non-neighboring regions; the smaller the geographic distance between regions, the more similar the law firm names; and law firm names within an economic region are typically more similar compared to the law firm names outside an economic region. On the other hand, the difference in the metrics of similarity is relatively small in each case, as is the correlation coefficient. It is therefore important to note that although the postulated phenomena do indeed exist, their magnitude is weak.

The three principal strategies for naming law firms are as follows: “Form”, “Law”, and “Service”. The first of these represents a minority approach, while the remaining majority is divided between two other strategies. In the European South of Russia and in the Volga economic region, “Service” is a dominant strategy. However, no other spatial patterns in law firms naming have been identified. Economic regions are typically characterized by a mixture of naming strategies, except for Volga-Vyatka, where the “Law” strategy prevails, and Volga and North Caucasus, where the “Service” strategy is dominant.

The contribution of this study is primarily theoretical and methodological in nature. It underscores the significance of firm names as a source of information for business- and economy-related research. It also proposes a technique for comprehensive analysis of branding in a geographical context through the use of modern natural language processing (NLP) solutions based on artificial neural networks. Additionally, it advocates for the utilization of disaggregated administrative open data for a spatial analysis. The findings demonstrate the effect of geographical proximity, neighboring, and economic relations on the similarity of the economic and, to a certain extent, cultural properties of the territories represented in firm names. The results generally align with the widely accepted classification of economic regions in Russia, although they highlight a few instances where a theoretically undesired internal heterogeneity within an economic region is observed. Finally, the study provides an illustration of a semi-automated brand classification within a specific economic sector and discovers a geographical variation in naming strategies that could be further investigated.

<div custom-style="References-title">References</div>

<div id="refs"></div>